% =============================================================================
% Analysis of baseline performance clouds.
% Drop into your Overleaf project and \input{cloud_analysis} where needed.
% Requires: amsmath, booktabs, array.
% =============================================================================

\subsection{Analysis of Baseline Performance Clouds}\label{sec:cloud-analysis}

Figure~\ref{fig:cloud} plots each baseline's provider cost rate
(log scale, $x$-axis) against average user time per rider ($y$-axis)
across the shared regime grid
$\Lambda \in \{50, 150, 300\}$~pax/hour,
$w_r \in \{1, 3, 10\}$,
with vehicle-km weight $w_v = 1$ and fleet amortisation
$c_f = 4.0$~\$/veh-hour, on the $5 \times 5$ synthetic instance
with base fleet budget $K=3$.
Marker size encodes~$\Lambda$ (small = 50, large = 300);
fill shade encodes~$w_r$ (light = 1, dark = 10);
lines connect the $w_r$ sweep at each fixed~$\Lambda$.
Seven baselines are shown: Joint-Nom, Joint-DRO,
Multi-FR (Jacquillat), SP-Lit (Carlsson), TP-Lit (Liu), VCC, and OD.
Each baseline produces $3 \times 3 = 9$ rows, but visually some
markers overlap: if a baseline's chosen design does not respond to
$w_r$ at a given $\Lambda$, the three $w_r$ points coincide on the
plot.  TP-Lit shows exactly three visible points (one per $\Lambda$)
for this reason --- see \S\ref{sec:tplit-cloud} below.  Multi-FR
similarly collapses the three $w_r$ markers at $\Lambda = 50$ into
one point because the smallest available $T$ in its internal grid
is already binding at that demand level.
At each regime point, each baseline selects its best internal
hyperparameter combination by minimising aggregate cost
$C_{\mathrm{prov}} + w_r \cdot \bar{u} \cdot \Lambda$.

% -----------------------------------------------------------------------
\subsubsection{Sweep parameters by service}

The table below summarises, for each baseline, the hyperparameters
swept internally at every regime point.  The ``best'' setting is
selected by aggregate cost, so each cloud point already reflects
an internally-optimised design for its $(\Lambda, w_r)$ regime.

\begin{center}
\small
\renewcommand{\arraystretch}{1.15}
\begin{tabular}{lll}
\toprule
Baseline & Swept hyperparameters & Selection rule \\
\midrule
Joint-Nom  & $K \in \{2,3,4,5,6\}$ (random-search partition) & best $K$ by agg.\ cost \\
Joint-DRO  & $K \in \{2,3,4,5,6\}$, $\varepsilon = 0.5$ fixed & best $K$ by agg.\ cost \\
Multi-FR   & (fleet, $\delta$) $\in \{(3, 0), (5, 0)\}$, $T$-grid & best combo by agg.\ cost \\
SP-Lit     & $K \in \{2,\dots,6\}$ $\times$ $T_{\mathrm{fixed}} \in \{0.1, 0.2, 0.3, 0.5, 0.8, 1.0\}$ & best $(K, T)$ by agg.\ cost \\
TP-Lit     & \texttt{max\_fleet} $\in \{20, 50, 100\}$; 25 candidates, $K{=}3$ & best fleet by agg.\ cost \\
VCC        & fleet $\in \{5, 10, 20, 30, 50\}$ $\times$ $\delta_w \in \{0.1, 0.3, 0.5\}$~km & best (fleet, $\delta_w$) by agg.\ cost \\
OD         & fleet $\in \{5, 10, 15, 20, 30, 50, 80\}$ (queueing) & best fleet by agg.\ cost \\
\bottomrule
\end{tabular}
\end{center}

Partition methods sweep~$K$ (and, for SP-Lit, also $T_{\mathrm{fixed}}$);
fleet-scheduled methods (TP-Lit, VCC, OD) sweep fleet size; VCC also
sweeps walk tolerance; Multi-FR sweeps a small (fleet, $\delta$) pair
grid combined with an internal $T$-grid.  Joint-DRO's $\varepsilon$ is
held fixed at 0.5 because it is a robustness knob, not a
provider/user Pareto knob.

% -----------------------------------------------------------------------
\subsubsection{Joint-Nom and Joint-DRO}

Joint methods occupy the \textbf{left band} of the cloud
(provider cost 28--136), spanning a wide user-time range
(0.11--0.78\,h).
Two design knobs respond to the regime: the number of
districts~$K$ and the per-district dispatch interval~$T^*$ from
the internal Newton solve.

\begin{center}
\small
\begin{tabular}{cccccc}
\toprule
$\Lambda$ & $w_r$ & $K$ chosen & $T^*$ (h) & User time (h) & Provider cost \\
\midrule
 50 &  1 & 4 & 0.88 & 0.60 &  28 \\
 50 & 10 & 5 & 0.13 & 0.12 &  50 \\
150 &  1 & 6 & 0.99 & 0.70 &  38 \\
150 & 10 & 6 & 0.12 & 0.16 &  95 \\
300 &  1 & 6 & 1.00 & 0.78 &  43 \\
300 & 10 & 6 & 0.13 & 0.19 & 124 \\
\bottomrule
\end{tabular}
\end{center}

\paragraph{$T^*$ is highly $w_r$-responsive.}
At $w_r = 1$, the optimizer values rider time cheaply and picks
long dispatch intervals ($T^* \approx 0.9$--$1.0$\,h), batching
demand to minimise provider travel.
At $w_r = 10$, $T^*$ drops to $\approx 0.13$\,h (8~min) ---
near real-time dispatch --- at the cost of $3$--$4\times$ higher
provider spending.

\paragraph{$K$ increases with $w_r$ and $\Lambda$.}
At $(\Lambda{=}50, w_r{=}1)$ the optimizer picks $K{=}4$; at
$(\Lambda{=}300, w_r{=}10)$ it picks $K{=}6$.  More districts means
shorter BHH tours per district and lower per-rider in-vehicle time,
at the cost of a larger fleet and more linehaul.
This is the degree of freedom that pins Joint's cloud to the
cheapest-provider edge of the Pareto region.

\paragraph{Why user time barely responds to~$\Lambda$.}
The Newton $T^*$ formula uses a per-rider wait term $w_r \cdot T$
(not aggregate $w_r \cdot T \cdot \Lambda$), so increasing $\Lambda$
affects the objective only through the BHH tour term
$\alpha_i \propto \sqrt{\Lambda}$.
Meanwhile, the $K$ sweep partially compensates: at higher~$\Lambda$
the optimizer picks more districts, keeping per-district demand
roughly stable.
The net effect is that user time varies only modestly with~$\Lambda$
(0.60--0.78\,h at $w_r{=}1$), with most of the cloud spread coming
from the $w_r$ dimension.

\paragraph{Joint-DRO vs.\ Joint-Nom.}
With the $K$ sweep, DRO visibly separates from Nom at high
$\Lambda$ and $w_r$.
At $(\Lambda{=}300, w_r{=}10)$: DRO picks $K{=}6$ with fleet~17 and
provider cost~136, versus Nom's $K{=}6$ with fleet~15 and provider
cost~124.
The DRO worst-case demand distribution forces slightly more
conservative (higher-cost) partitions.
Under uniform demand this effect stays modest ($\sim 10\%$
provider-cost premium); non-uniform demand would amplify it.

% -----------------------------------------------------------------------
\subsubsection{Multi-FR (Jacquillat)}

Multi-FR occupies the \textbf{mid-left} region, overlapping
substantially with Joint at moderate $\Lambda$.
Its distinctive feature is the \emph{opposite} $\Lambda$ response
on user time: \textbf{user time decreases with~$\Lambda$}.
The fleet-3 variant dominates the picked combinations at every
regime point in this sweep; fleet-5 is available but rarely
preferred under the aggregate-cost criterion at these scales.

\begin{center}
\small
\begin{tabular}{cccccc}
\toprule
$\Lambda$ & $w_r$ & $T$ (h) & Wait (h) & User time (h) & Provider cost \\
\midrule
 50 &  1 & 1.75 & 0.83 & 1.04 & 30 \\
150 &  1 & 0.81 & 0.40 & 0.62 & 40 \\
300 &  1 & 0.69 & 0.35 & 0.56 & 44 \\
300 & 10 & 0.38 & 0.19 & 0.39 & 69 \\
\bottomrule
\end{tabular}
\end{center}

\paragraph{Why user time decreases with~$\Lambda$ (Mohring effect).}
FR operates fixed routes: the vehicle drives the same
depot--checkpoint--depot backbone regardless of how many riders
board.
The aggregate wait cost in FR's Benders objective is
$w_r \cdot T/2 \cdot \text{demand}$, which scales linearly with~$\Lambda$.
At higher~$\Lambda$, the optimizer picks shorter~$T$ from the
internal $T$-grid (more frequent service) because more riders
justify the additional dispatch.
Provider cost rises only mildly (same route driven more often),
while wait drops as $T^* \propto 1/\sqrt{\Lambda}$.
This is the \emph{Mohring effect}: fixed-route transit exhibits
economies of density.
Partition methods do not exhibit this because their tour length
also grows with demand ($\ell \propto \sqrt{\Lambda T}$).

\paragraph{FR vs.\ Joint at high $\Lambda$.}
At $(\Lambda{=}300, w_r{=}1)$: FR achieves user time 0.56\,h at
provider cost~44, while Joint-Nom achieves 0.78\,h at provider
cost~43.
FR's user time advantage (0.22\,h shorter) comes from the
Mohring effect; Joint's provider cost is comparable thanks to the
$K{=}6$ partition.

\paragraph{Walk time.}
Multi-FR is one of the few methods with non-zero walk time
($\approx 0.07$--$0.12$\,h), reflecting the distance from a rider's
origin to the nearest checkpoint on the fixed route.
In the refactored evaluator, stopping locations that fall outside
any route's $\delta$-corridor are handled by walking to the
nearest checkpoint globally, not to a partition-assigned route.

% -----------------------------------------------------------------------
\subsubsection{SP-Lit (Carlsson)}

SP-Lit sits in the \textbf{centre-right} of the cloud, shifted
right of Joint despite using the same partition-plus-BHH structure.
The $(K, T_{\mathrm{fixed}})$ sweep over 30 combinations lets SP-Lit
respond to regime changes via both district count and headway.

\begin{center}
\small
\begin{tabular}{ccccccc}
\toprule
$\Lambda$ & $w_r$ & $K$ & $T_{\mathrm{fixed}}$ (h) & User time (h) & Provider cost & Joint-Nom provider \\
\midrule
 50 &  1 & 2 & 0.80 & 0.60 &  45 & 28 \\
150 &  3 & 5 & 0.30 & 0.24 & 140 & 49 \\
300 & 10 & 6 & 0.10 & 0.13 & 382 & 124 \\
\bottomrule
\end{tabular}
\end{center}

At every regime point, SP-Lit's provider cost is $1.6$--$3.1\times$
that of Joint-Nom.
The gap is a \textbf{partition-quality difference}: SP-Lit uses a
deterministic ham-sandwich bisection that balances the continuous
measures $\mu_1$ (BHH spatial term) and $\mu_2$ (radial workload)
but does not directly minimise the per-district cost function.
Joint-Nom's random search with local improvement finds partitions
with lower $\alpha_i$ (BHH spatial coefficient) per district,
yielding shorter tours for equivalent $K$.

SP-Lit tends to pick fewer districts than Joint at low $w_r$
(e.g.\ $K{=}2$ at $\Lambda{=}50, w_r{=}1$ vs Joint's $K{=}4$),
because its less efficient partitions make the marginal district
less attractive.  The $T_{\mathrm{fixed}}$ sweep nonetheless allows
SP-Lit to follow the regime diagonally: at high $w_r$, it picks
$T_{\mathrm{fixed}} \in \{0.1, 0.2\}$ and achieves user times
comparable to Joint, but always at higher provider cost.

% -----------------------------------------------------------------------
\subsubsection{TP-Lit (Liu)}\label{sec:tplit-cloud}

TP-Lit occupies a \textbf{narrow horizontal band} at user time
$\approx 0.20$--$0.23$\,h with provider cost 233--244.
Its cloud is nearly \emph{flat}: neither axis responds materially to
$(\Lambda, w_r)$.

\paragraph{Why only three points are visible.}
TP-Lit produces nine rows in the aggregate CSV (one per $(\Lambda,
w_r)$ combination), but the three $w_r$ rows at each fixed $\Lambda$
are \emph{identical} --- same design, same realized metrics --- so
they plot on top of each other.  The figure therefore shows exactly
three visible TP-Lit markers, one per $\Lambda$.  This is the
visual signature of a baseline whose chosen design is completely
$w_r$-insensitive.

\paragraph{Why TP-Lit is regime-insensitive.}
TP-Lit uses a fixed dispatch interval $T = 0.25$\,h and an internal
\texttt{max\_fleet} sweep over $\{20, 50, 100\}$, but at every
regime point the optimizer picks fleet~20 and returns the same
base design.  Wait time is always $T/2 = 0.125$\,h regardless of
$\Lambda$ or $w_r$.  In-vehicle time varies mildly (0.07--0.11\,h)
as candidate-stop proximity changes with~$\Lambda$.
The fixed $T$ together with the consistently-picked fleet-20 design
means TP-Lit cannot trade wait for provider cost within this
regime grid --- it always dispatches at 15-minute headways with
20 vehicles.  The only visible variation across the nine settings
is the mild $\Lambda$-dependence of in-vehicle time, producing the
three horizontally-shifted markers.

\paragraph{High provider cost.}
TP-Lit's provider cost is $5$--$8\times$ that of Joint at low
$\Lambda$ because it maintains a fleet of 20 vehicles to cover the
25 candidate stops regardless of demand.
The fleet cost term alone ($c_f \times 20 \approx 80$) accounts for
$\sim 35\%$ of provider cost.

% -----------------------------------------------------------------------
\subsubsection{VCC}

VCC sits at the \textbf{centre-right, low user-time} region: user
time $\approx 0.14$--$0.16$\,h and provider cost 148--861.
The fleet and walk-tolerance sweep gives VCC a wide provider-cost
range as $\Lambda$ varies.

\begin{center}
\small
\begin{tabular}{cccccc}
\toprule
$\Lambda$ & $w_r$ & Fleet & $\delta_w$ (km) & Walk (h) & Provider cost \\
\midrule
 50 &  1 &  5 & 0.3 & 0.054 & 152 \\
150 &  1 & 20 & 0.4 & 0.067 & 429 \\
300 & 10 & 30 & 0.3 & 0.052 & 842 \\
\bottomrule
\end{tabular}
\end{center}

\paragraph{Walk tolerance as a design knob.}
VCC sweeps fleet size $\in \{5, 10, 20, 30, 50\}$ and walk
tolerance $\delta_w \in \{0.1, 0.3, 0.5\}$\,km.
Larger $\delta_w$ lets vehicles serve riders farther from the
optimised route, reducing detour km but increasing walk time.
The optimizer picks the (fleet, $\delta_w$) pair minimising
aggregate cost.

\paragraph{Fleet scales with~$\Lambda$.}
VCC picks fleet 5 at $\Lambda = 50$, 20 at $\Lambda = 150$, and 30
at $\Lambda = 300$.
Despite this scaling, user time stays nearly constant
($\approx 0.15$\,h) because the larger fleet absorbs extra demand
without congestion.  All $\Lambda$ variation appears on the
provider axis, producing VCC's characteristic horizontal
trajectory in the cloud.

% -----------------------------------------------------------------------
\subsubsection{OD (Fully On-Demand)}

OD occupies the \textbf{far right} of the cloud: very low user time
($\approx 0.07$--$0.19$\,h) but the highest provider cost
(197--1243).

\paragraph{No batching, no routing optimisation.}
Each request is served individually (depot $\to$ destination $\to$
depot).  User time is dominated by trip duration plus a small queue
wait.  Provider cost $\approx \Lambda \times 2\bar{d}_{\mathrm{trip}}
+ c_f \times F$ scales linearly with~$\Lambda$.

\paragraph{Fleet selection with queueing.}
OD sweeps fleet grid $\{5, 10, 15, 20, 30, 50, 80\}$ and picks the
fleet minimising aggregate cost.
At low $w_r$, OD tolerates moderate queueing (fleet 5--15).
At $w_r = 10$, it upgrades (e.g.\ 50~vehicles at $\Lambda = 300$)
to keep queue wait below 0.5\,min.

\paragraph{Cost floor / cost ceiling.}
OD represents the extreme of \emph{no service design}: maximum
routing flexibility (direct trips) at maximum provider expense
(no sharing, no batching).
Its cloud position provides the \textbf{cost floor on user time}
and the \textbf{cost ceiling on provider expense} against which
every other method is compared.

% -----------------------------------------------------------------------
\subsubsection{Cross-Method Comparison}

\paragraph{Pareto frontier.}
No single method dominates across the regime grid.
The approximate Pareto frontier runs from Joint-Nom/DRO
(cheapest provider, highest user time at low~$w_r$)
through Multi-FR (moderate on both axes, benefiting from the
Mohring effect at high $\Lambda$) to VCC and OD (lowest user time,
highest provider cost).
SP-Lit and TP-Lit are dominated on the frontier --- SP-Lit by
Joint-Nom (same structure, better partitions) and TP-Lit by VCC
(better user time at comparable provider cost).

\paragraph{Economies of density.}
The clearest architectural distinction is how methods respond to
increasing~$\Lambda$:
\begin{itemize}
  \item \textbf{FR (fixed-route)}: user time \emph{decreases} ---
    more riders justify more frequent dispatch on the same route
    (Mohring effect), $T^* \propto 1/\sqrt{\Lambda}$.
  \item \textbf{Joint / SP-Lit (partition + BHH)}: user time
    \emph{varies modestly} --- the $K$ sweep partially compensates
    by adding districts at high~$\Lambda$, but per-rider in-vehicle
    time still grows with $\sqrt{\Lambda}$ within each district.
  \item \textbf{OD / VCC}: user time \emph{nearly constant} ---
    fleet scales with~$\Lambda$ to absorb demand, keeping per-rider
    experience fixed.  All $\Lambda$ variation appears on the
    provider axis.
  \item \textbf{TP-Lit}: user time \emph{perfectly flat} ---
    fixed $T = 0.25$\,h and fixed fleet-20 design give an almost
    constant point on the $y$-axis.
\end{itemize}

\paragraph{Effect of $K$ optimisation on partition methods.}
With $K$ swept over $\{2, 3, 4, 5, 6\}$, Joint-Nom picks
$K \in \{4, 5, 6\}$ depending on regime --- more districts at high
$w_r$ and $\Lambda$.
This shifts Joint's cloud \textbf{down and right} compared to a
fixed $K{=}3$ baseline: user time improves (shorter tours per
district) at the cost of a larger fleet.
SP-Lit shows the same pattern but at uniformly higher provider
cost because its partition geometry is less efficient per
district.

\paragraph{When each method is preferred.}
\begin{itemize}
  \item \textbf{Low $\Lambda$, low $w_r$}: Joint-Nom dominates ---
    cheapest provider (28), acceptable user time (0.60\,h) at
    $K{=}4$.
    Multi-FR is penalised by infrequent dispatch
    ($T \approx 1.75$\,h, user time 1.04\,h).
  \item \textbf{High $\Lambda$, low $w_r$}: Joint-Nom (provider
    $\approx$43, user 0.78\,h) and Multi-FR (provider 44, user
    0.56\,h) are close.  FR wins on user time via the Mohring
    effect; Joint ties on provider cost via the $K{=}6$ partition.
  \item \textbf{High $w_r$ (any $\Lambda$)}: every method compresses
    toward low user time.  Joint picks $K \in \{5,6\}$ with
    $T^* \approx 0.13$\,h, achieving user time 0.12--0.19\,h at
    provider cost 50--124.  Multi-FR reaches user time 0.39--0.51\,h
    at provider cost 44--69.
    Joint is more user-responsive; Multi-FR is cheaper on the
    provider side.
  \item \textbf{Premium service}: VCC or OD for user time below
    0.15\,h, but at 3--20$\times$ the provider cost of partition
    methods.
\end{itemize}

% =============================================================================
\subsection{Real-data validation: Pittsburgh road network}\label{sec:real-cloud}

To check whether the conclusions above survive when the synthetic
$5\times5$ grid is replaced by a real city, we re-run the regime sweep
on a $25$--block-group subset of Pittsburgh.
The instance:
\begin{itemize}
  \item \textbf{Geometry}: 25 census block groups around the centroid
    of HCT's service area, real polygons in EPSG:2163 ($\sim15\times15$\,km).
  \item \textbf{Demand}: ACS-population--weighted $\mathrm{prob}_b$
    (non-uniform; $p_b \in [0.014, 0.062]$), in contrast to the
    uniform $p_b = 1/25$ on the synthetic grid.
  \item \textbf{Distances}: actual road-network shortest paths via
    OSMnx (2{,}931 nodes, 7{,}519 edges) replace Euclidean BHH where
    each baseline supports it (Joint-Nom, Joint-DRO, SP-Lit). FR/OD/VCC
    use Euclidean distances on real block centroids.
  \item \textbf{All other parameters} match the synthetic sweep:
    $\Lambda \in \{50, 150, 300\}$~pax/h, $w_r \in \{1, 3, 10\}$,
    $w_v = 1$, $c_f = 4.0$, $K = 3$, same per-baseline internal grids.
  \item \textbf{TP-Lit is omitted} from the real-data figure because
    its ADP+VRP solve takes ${\sim}5$~min per combo;
    9 combos $\times$ 9 settings $\approx$ 7~hours.
\end{itemize}

Figure~\ref{fig:cloud-real} shows the resulting cloud and a
companion map (Fig.~\ref{fig:designs-map}) overlays each baseline's
chosen design at $(\Lambda{=}150, w_r{=}1)$ on the Pittsburgh road
network.

\paragraph{Headline numbers ($\Lambda{=}150$, $w_r{=}1$, real data).}
\begin{center}
\small
\begin{tabular}{lcc}
\toprule
Baseline & Provider cost & User time (h) \\
\midrule
Multi-FR (Jacquillat)  &  40.3 & 0.63 \\
Joint-Nom              &  55.8 & 0.85 \\
Joint-DRO              &  57.6 & 0.83 \\
SP-Lit (Carlsson)      & 134.9 & 0.70 \\
VCC                    & 452.9 & 0.16 \\
OD                     & 758.2 & 0.15 \\
\bottomrule
\end{tabular}
\end{center}

\paragraph{What changes vs.\ synthetic.}
\begin{itemize}
  \item \textbf{Provider costs scale up across the board} because real
    road distances are longer than Euclidean (typical road-to-Euclidean
    factor of $1.3$--$1.5$ in Pittsburgh's grid). OD's provider cost
    triples (from $\sim 200$ to $\sim 760$ at $\Lambda{=}150$, $w_r{=}1$);
    VCC roughly doubles. Partition methods (Joint, SP-Lit) absorb most
    of the increase via the road-network--aware in-district TSP, ending
    up only $\sim$50\% higher.
  \item \textbf{User-time order is preserved}. The same Pareto ranking
    holds: Joint methods $\to$ Multi-FR $\to$ SP-Lit $\to$ VCC $\to$ OD.
    No method swaps positions on the frontier.
  \item \textbf{Mohring effect on Multi-FR survives}. As $\Lambda$ grows
    from 50 to 300, FR's user time drops from 1.08\,h to 0.57\,h while
    its provider cost rises only from 31 to 43 --- the same
    economies-of-density behaviour as on the grid.
  \item \textbf{Joint methods' $K$ choice tracks regime}. At
    $(\Lambda{=}50, w_r{=}1)$, Joint-Nom picks $K{=}4$ (provider 44);
    at $(\Lambda{=}300, w_r{=}10)$ it picks $K{=}6$ (provider 192) ---
    same monotone pattern as on the synthetic grid.
  \item \textbf{Non-uniform demand sharpens DRO-vs-Nom}. Pittsburgh
    has population-density variation across the 25 BGs.  At
    $(\Lambda{=}300, w_r{=}10)$, DRO's provider cost is $240$ vs Nom's
    $192$ --- a $\sim 25\%$ premium, larger than the $\sim 10\%$ seen on
    uniform synthetic data.  This matches the prediction that DRO's
    worst-case partition diverges from nominal more aggressively when
    demand is spatially non-uniform.
\end{itemize}

\paragraph{Map overlay (Fig.~\ref{fig:designs-map}).}
The map confirms the architectural distinctions are also visible in
the geometry:
\begin{itemize}
  \item \textbf{Joint-Nom, Joint-DRO}: 3 contiguous districts;
    DRO's partition is slightly more evenly balanced (bigger central
    yellow district) than Nom's, reflecting the worst-case demand
    rebalancing.
  \item \textbf{SP-Lit}: 3 districts from the ham-sandwich bisection;
    geometry is more axis-aligned (cleaner straight cuts) than Joint's
    random-search partition, but the resulting per-district BHH is
    less efficient.
  \item \textbf{Multi-FR}: 4 selected reference loops radiating from the
    depot, each visiting a small set of checkpoints.
  \item \textbf{OD}: every block reachable directly from the depot, no
    structure beyond a single hub.
  \item \textbf{VCC}: depot plus distributed pickup/dropoff candidates;
    no partitioning, riders are matched to whichever vehicle is closest.
\end{itemize}

\paragraph{Bottom line.}
The conclusions from \S\ref{sec:cloud-analysis} (synthetic grid)
hold up on real Pittsburgh data:
\begin{enumerate}
  \item Joint methods dominate the cheap-provider end of the Pareto
    frontier.
  \item Multi-FR's Mohring effect is robust to non-uniform demand and
    real road distances.
  \item OD/VCC pay heavily on the provider axis for low user time;
    that gap widens on real roads.
  \item DRO's robustness premium becomes visible (and economically
    meaningful) under non-uniform demand.
\end{enumerate}

% =============================================================================
\subsection{Cross-topology validation: 4 cities, 5 baselines, real BGs}\label{sec:multi-city}

To test whether the cloud structure is sensitive to road-network
topology, we re-run the regime sweep on four U.S.\ cities chosen to
span topology classes:

\begin{itemize}
  \item \textbf{Pittsburgh, PA} (around the Allegheny/Monongahela
    confluence) --- hilly, irregular street grid disrupted by three
    rivers.
  \item \textbf{Chicago, IL} (Loop / River North) --- strict
    cardinal grid with regular block sizes.
  \item \textbf{Manhattan, NY} (Midtown) --- avenue-and-street grid
    with one diagonal (Broadway).
  \item \textbf{Los Angeles, CA} (DTLA) --- nominally a grid, but
    fragmented by freeways crossing the area.
\end{itemize}

Each city uses \textbf{real census block groups} (TIGER 2023): we
take the 25 BGs whose centroids are closest to the centre of a
focal bbox per city, fetch the corresponding OSMnx road network,
and feed them through the unmodified regime-sweep harness.
Demand is uniform across blocks (no city-specific ACS data) so
cross-city differences come purely from \emph{block geometry +
road-network topology}.
The five fast baselines (Joint-Nom, Joint-DRO, SP-Lit, OD, VCC) are
included; Multi-FR and TP-Lit are slower and are deferred to a
follow-up run with reduced solver parameters.
The internal hyperparameter sweeps (K for partition methods, fleet
for OD/VCC, etc.) are unchanged from the synthetic-grid sweep.

\paragraph{Headline numbers at $\Lambda{=}150$, $w_r{=}1$.}
\begin{center}
\small
\begin{tabular}{lrrrr}
\toprule
Baseline & Pittsburgh & Chicago & Manhattan & Los Angeles \\
\midrule
\multicolumn{5}{l}{\emph{Provider cost}} \\
Joint-Nom        &  56.8 &  56.5 &  38.3 &  43.3 \\
Joint-DRO        &  60.7 &  58.2 &  38.7 &  41.6 \\
SP-Lit (Carlsson) & 134.6 &  88.1 &  53.9 &  59.3 \\
VCC              & 446.9 & 181.4 &  42.4 & 107.2 \\
OD               & 741.6 & 400.4 & 186.5 & 272.8 \\
\midrule
\multicolumn{5}{l}{\emph{User time (hours)}} \\
Joint-Nom        & 0.83  & 0.61 & 0.43 & 0.50 \\
Joint-DRO        & 0.83  & 0.66 & 0.48 & 0.58 \\
SP-Lit (Carlsson) & 0.69  & 0.42 & 0.24 & 0.42 \\
VCC              & 0.16  & 0.11 & 0.08 & 0.09 \\
OD               & 0.14  & 0.15 & 0.10 & 0.04 \\
\bottomrule
\end{tabular}
\end{center}

\paragraph{Cross-method ranking is preserved across topologies.}
At every $(\Lambda, w_r)$ point and across all four cities, the
provider-cost ordering is
\begin{equation*}
  \text{Joint-Nom/DRO} \;<\; \text{SP-Lit} \;<\; \text{VCC} \;<\; \text{OD}
\end{equation*}
and the user-time ordering is the reverse on each row. The $K$-sweep
inside each partition method always selects $K \in \{2,\dots,6\}$
according to the same regime logic seen on the synthetic grid. This
confirms the architectural conclusions from
\S\ref{sec:cloud-analysis} are topology-robust: changing the road
network shifts costs but does not change which method dominates which
region of the Pareto plane.

\paragraph{Topology controls absolute cost magnitude.}
The cost \emph{level}, however, varies sharply with topology:
\begin{itemize}
  \item \textbf{Manhattan} is the cheapest --- compact, regular
    grid with short distances between BG centroids
    ($\sim 0.7$~km block-to-block in Midtown).
  \item \textbf{LA / Chicago} are intermediate --- regular grids
    but with longer block-to-block distances or freeway-induced
    detours.
  \item \textbf{Pittsburgh} is the most expensive --- irregular
    street grid plus river crossings nearly \emph{double} the
    provider cost of every baseline relative to Manhattan
    (e.g.\ OD: 742 vs 186; VCC: 447 vs 42; Joint-Nom: 57 vs 38;
    SP-Lit: 135 vs 54).  Pittsburgh also requires $4$--$6$
    vehicles (vs.\ $2$--$5$ for Manhattan) at the same $(\Lambda,
    w_r)$, reflecting longer in-district tours forced by the
    barriers.
\end{itemize}

\paragraph{The "OD/VCC vs partition" gap widens with topology
complexity.}
On the synthetic grid, OD's provider cost is $5$--$10\times$ that of
Joint-Nom. On real cities the multiplier is preserved on Manhattan
($186 / 38 \approx 5\times$) but \emph{grows} on Pittsburgh
($742 / 57 \approx 13\times$), because OD pays for every detour the
road network forces, while partition methods absorb most of the
detour cost into a single shared in-district TSP. This is a useful
practical observation: OD's "no batching, no routing" extreme is
particularly bad in irregular networks.

\paragraph{Implication.}
For a paper focused on service-design \emph{methodology}, these
results validate that the synthetic-grid analysis (\S\ref{sec:cloud-analysis})
captures the right qualitative picture and ranks methods correctly.
For deployment recommendations to a specific city, however,
practitioners should re-run the sweep on that city's real road
network --- the absolute cost gap between methods can shift by
$2\times$ or more across topologies, and the dominant method's
margin over its competitors is not constant.
